---
title: 'Homework #1'
author: "M. Drew LaMar"
date: "Due: Wednesday, September 12, 2018, 1:00 pm"
output:
  word_document: default
  html_document:
    toc: yes
    toc_depth: 3
---

## Chapter 1

### Chapter 1: #14

 (a) Answer to (a)
 (b) Answer to (b)
 (c) Answer to (c)

### Chapter 1: #16

### Chapter 1: #17

### Chapter 1: #18

### Chapter 1: #19

### Chapter 1: #20

### Chapter 1: #24

---

## Chapter 2

### Chapter 2: Practice Problem #12 (example - please delete me!)

 (a) What type of table is this?
 
Contingency table
 
 (b) Display these same data in a mosaic plot.
 
First, let's read in the data and take a quick peek.
 
```{r}
mydata <- read.csv("http://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q12ConvictionsAndIncome.csv")
head(mydata)
str(mydata)
```
First question to ask yourself is:  **What are the variables and their types?**  It looks like we have two categorical variables: `hasConvictions`, with levels `yes` and `no`, and `incomeLevel`, with levels `comfortable`, `adequate` and `inadequate`.

Second question to ask yourself is:  **Is the data in processed or raw form?**  In other words, *are rows objects from a population, and are columns variables/measurements on those objects?*  In this case, it appears not.  It looks like someone already computed the frequencies of occurrences of the different categorical variables.  Since we know we want to do a mosaic plot, we need to make sure the data is in the right form to send to the function `mosaicplot`.  The [Chapter 2 R-Code Examples](http://rstudio-pubs-static.s3.amazonaws.com/150233_15c3dbafd8f34216901bded8d8e299d0.html#example-2.3a.-bird-malaria) show us that we need to have the data in **contingency table form**.  *This is one of our first examples of why you need to learn how to change the form of your data!!!*  Let the data munging begin!!!

In this case, we'll just do this manually and learn some more sophisticated tools later in the course.  Since you've learned about the `matrix` data type, let's use that (a matrix is another form of a table).  We'll put `hasConvictions` as rows and `incomeLevel` as the columns:

```{r}
(mytable <- matrix(mydata$frequency, byrow = TRUE, nrow = 2))
```
Looks good, except names for rows and columns would be good.  **Make sure you name them in the correct order!!!**
```{r}
rownames(mytable) <- c("yes", "no")
colnames(mytable) <- c("comfortable", "adequate", "inadequate")
mytable
```
Now just in case you like to have `incomeLevel` increasing from left to right, we can reorder the columns.  Note that while `1:3` creates a vector (`r 1:3`), we can also count backwards with the command `3:1`, giving (`r 3:1`).
```{r}
(mytable <- mytable[,3:1])
```
That's better!  Now, our mosaicplot - we'll need to transpose the matrix to make sure the explanatory variable (`incomeLevel`) is along the horizontal axis.
```{r}
mosaicplot(t(mytable), 
           xlab = "Income Level",
           ylab = "Convictions",
           main = "")
```

Yay!!!

 (c) Categorical (ordered). Groups should be
arranged by increasing income.
 
 (d) The relative frequency of convictions decreases as available income increases.
 
 (e) The mosaic plot, which made it easier to see the pattern. Whereas the table gives the frequencies, the graph visualizes the association between the variables.
 
### Chapter 2: #20

### Chapter 2: #23

### Chapter 2: #29

### Chapter 2: #32

### Chapter 2: #33

### Chapter 2: #34

### Chapter 2: #35

### Chapter 2: #36

## More examples of R-Markdown awesomeness (please delete me!)

This is regular text.  You can even do math!!!  This is **bold**.

$$
a^2 + b^2 = c^2
$$

Use a single dollar sign to do math "inline", like so:  $\alpha = \beta$.

> This is a blockquote to your teaching assistants.  Use this to leave them a specific comment about your work (for example, "I'm having trouble with this.  Feedback appreciated.")

Here's an example of embedded R code:

```{r}
x <- 4
(y <- x + 2)
```

You can even embed R code "inline", like math.  So, for example, I can say y = `r y`, or even embed it into math:  $y = `r y`$.  If I change the code in the R chunk above, it will modify the text accordingly.